Episodic memory: mental time travel or a quantum `memory wave' function?

### 2019-01-11 main.tex
fleshing out a bunch of stuff in the paper; one more section to go, on extensions to non-temporal domains

One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.  An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.  Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.  This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12, AlyEtal18}.  Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}. This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11}. Recent work also suggests that the lateral entorhinal cortex plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.  These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14}.  In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.  Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.  First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).  Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.  These reactivations explain how our brains reactivate prior contexts when we remember the past.

Although reactivating context is often \textit{described} as a sort of mental time travel back to the one specific moment being remembered, there is some apparent inconsistency between this description and how this process is actually characterized mathematically or measured neurally.  In particular, both the theory and neural measures described in the preceding paragraphs characterize episodic recall mathematically as reflecting the reactivation of a weighted blend of thoughts from a \textit{range} of previously experienced times-- not to any one \textit{specific} time.  This is not simply a matter of imprecision in mental time travel (i.e., a reflection of the low effective resolution with which we can revisit specific moments from our past).  Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences. The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each drifting at a different timescale.  The notion of thinking of any one particular moment from the past is incompatible with this multiple timescales view, in that our thoughts and brain activity patterns reflect information from a range of moments, even when we are not specifically engaged in remembering.  If what we consider to be a ``moment'' or ``now'' is guided by our neural representations of our ongoing experiences, then this implies we are continually experiencing \textit{many} ``nows.''  If our thoughts are continually spread across a range of times, the notion of mentally revisiting any particular moment loses its meaning.

Where the mental time travel framing breaks down most notably is when we consider scenarios in which the particular blend of timepoints reflected in one's current thoughts do not come from temporally contiguous events.  For example, understanding how a new experience fits in with our broader understanding might require integrating information gleaned from many experiences separated in time, each with their own contextual attributes.  That we can integrate information across experiences suggests that the integration of incoming information at different rates need not be a purely mechanical process, but rather may reflect a deeper functional role.  When we simultaneously reactivate thoughts related to separated moments from our past, this provides a means of associating and relating those temporally distinct experiences.  This is how we can integrate information across the \textit{many} discrete experiences relevant to ``now.''

@@

Event boundaries are one example of the broader temporal covariance structure that is characteristic of ongoing experience.  Whereas classic approaches to studying memory (e.g., list-learning paradigms and other trial-based experiments) encourage researchers to treat each moment, trial, or stimulus as separable from the rest of experience, context-based theories of memory (including situation models and event-based models) posit that each moment derives meaning through its deep ties to other experiences.  This rich tapestry of interacting moments and experiences forms the scaffolding for interpreting our new experiences and for retrieving information concerning our past.  According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e., how the event relates to the rest of ones experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!  This begs the question: which thoughts, from which times, do we reactivate when we remember our past experiences?

\subsection*{How is our past replayed when we remember?}
Our ongoing experiences can cue us to dredge up information about our past.  This can happen explicitly (e.g., during a conversation about a prior event) or implicitly (e.g., when something happening now reminds you of what happened before).  Context-based theories of memory posit that the particular memories that our ongoing experiences cue will be related~\citep[contextually, semantically, functionally, etc.; e.g., ][]{PolyEtal09}.  One potential challenge to studying how we remember the past relates to detangling which links between our experiences are specifically driven by our memory systems versus which are imposed ``externally'' through the similarity structure of those experiences.  For example, consider your daily commute into work.  If you were to verbally recount a specific day's commute later, one might expect some aspects of your description to apply to other days' commutes as well (e.g., other commutes where you took a similar route or travelled via similar modes of transportation, listened to similar music along the way, travelled at a similar time, etc.).  But to what extent could one hope to separate out descriptive matches driven solely by similarities in those commuting experiences, versus true reactivations (in memory) of specific memories of those other prior commutes?

Once insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \citep{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.  The authors used topic models~\citep{BleiEtal03} applied to human annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.  When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change. This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \citep{HeusEtal18c} perform detailed analyses of these putative events and how they are remembered.  Notably, the off-diagonal entries of the temporal correlation matrix (which reflects content similarity between temporally distant events in the episode) are nearly all close to 0.  This is surprising, because the episode itself comprises a murder mystery with recurring motifes that must be associated in the viewer's mind in order to make sense of the episode.  Nonetheless, when considering solely the semantic content of each moment of the episode, those long timescale associations are nearly entirely absent.  To the extent that this finding extends to real-world experiences, it comports with Heraclitus' notion that one cannot step in the same river twice~\citep{Hera}.

@@

Although the specific content of each moment of the episode is largely unique, participants' subjective experiences of the episode entail weaving together those temporally separated moments into a cohesive narrative.  This is reflected in how people verbally recount the episode later.  If we examine the temporal autocorrelation matrices of participants' recalls of the television episode (characterized using the same topic model used to capture the content of the episode; Figs.~\ref{fig:corrmats}B, C), we see a different structure than that of the episode itself.  Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off diagonal correlations between temporally distant events are substantially stronger in participants' recalls than in the original episode.  Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode and participants' recalls (Fig.~\ref{fig:reinstatement}A).  Whereas the autocorrelations for the original episode rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around 0.2.  These analyses indicate that the way participants \textit{recount} temporally separated events imposes additional similarity on those events, beyond what was present in the original events themselves.  One can then ask: is this simply a matter of imprecision in recall, either with respect to participants' recalls themselves, or with respect to the quality with which they are characterized via semantic feature vectors?  Or might there be a functional role to that additional imposed similarity structure?

To examine which specific events are described (during recall) as more similar than what their original experiences (as characterized by semantic feature vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.  Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were described in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).  The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has been murdered.  The language used by human annotators describing the moment-by-moment contents of the scene used a similar mix of themes only for moments in the episode that were temporally proximal to the event in question (gray line).  However, the way participants \textit{described} what happened when they recounted the episode later revealed a much different pattern.  Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering (Fig.~\ref{fig:reinstatement}C).  Further, the way they described this example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters.  Taken together, this suggests that these specific events were associated in participants' memories, perhaps contributing to (or reflecting) their understanding that those events were linked in the episode's narrative.

@@

A rich recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.  For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, were functionally associated in memory.  Nonetheless, a potential challenge to testing this hypothesis in the laboratory is that in most memory studies there is no particular reason for participants to functionally associate the stimuli-- there is no ``deeper narrative'' to make sense of.  However, random word list learning experiments might happen to contain (for a given list) semantically related words.  And indeed, a number of studies indicate that participants spontaneously form associations between semantically related words on random word lists, even when those related words are separated in time~\citep[e.g.,][]{WixtRohr94, MannKaha12, MannEtal12}.  This suggests that our memory systems pick up on statistical regularities, even in ostensibly ``random'' stimulus sequences, and that we leverage those regularities when we recall our past~\citep{PolyEtal09}.  Other work suggests that our memory systems may also leverage these regularities to parse our ongoing experiences into discrete events~\citep{SchaEtal13}.  Are these examples in the literature on random list learning studies, indicating that our memory systems leverage ``accidental'' structure in random sequences, a reflection of the same fundamental principle driving the sorts of memory reactivation function displayed in Figure~\ref{fig:reinstatement}B?

@@

\section*{The shapes of remembered experiences}
In a recent study~\citep{HeusEtal18c} my lab examined functional neuroimaging and behavioral data collected by \cite{ChenEtal17} as participants viewed and then verbally recalled an episode of a television show.  We used topic models~\citep{BleiEtal03} to characterize the moment-by-moment content of the episode and of participants' recalls.  Each moment was quantified via a \textit{topic vector} that reflected the mix of topics (themes) the model identified in that moment.  In this way, the sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.  The sequence of topic vectors for a given participant's recalls traces out an analogous trajectory.  This characterization provides a geometric framework for comparing the \textit{shapes} that reflect how naturalistic experiences unfold over time, to the shapes of how we remember those experiences later (Fig.~\ref{fig:trajectories}).  One can then use the geometric transformations needed to map an experience's trajectory onto the trajectory of its later remembering to characterize the quality and content of memory.

@@

The specific ties between unfolding events forms a scaffolding for understanding how those events are related.  In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\citep{JollChan18}.  \cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats}, \ref{fig:reinstatement}, and~\ref{fig:trajectories}, reveals another key property of how we remember.  Although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.  This has potentially exciting implications for how we communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.  It also has implications for how we learn by analogy: perhaps events that share schemas~\citep[e.g., ][]{BaldEtal18}  unfold in similar ways (e.g., have similar shapes).  Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).

@@

\subsection*{Extensions to spatial and semantic memory}
Just as we mentally spread ourselves across a distribution of times, we can perform similar mental operations when we conceptualize and retrieve other types of information.  For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12} or spatial~\citep{MillEtal13} content being mentally manipulated.  The prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.  Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.  Traditional approaches then ``clean up'' this distribution over possible mental states by taking its expected or maximum value.  However, following the ``quantum wave function'' logic, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.  What are the implications of this view?

\section*{Concluding remarks}

>>>>

One practical way to distinguish content from context representations is to separate mental (or neural) representations that drift rapidly \citep[content;][]{PolyEtal05, MannEtal12} or gradually \citep[context;][]{PolyEtal05, MannEtal11, HowaEtal12, LohnEtal18, LongKaha18, FolkEtal18}.  An intuition for why contextual representation might be expected to evolve gradually is laid out by \cite{PolyKaha08}.  Essentially, slowly drifting thoughts may be used to ``index'' more rapidly drifting thoughts that are temporally proximal.  This notion, that our brain maintains parallel mental representations drifting at different time scales, has been well-characterized in a series of elegant fMRI and ECoG studies by Uri Hasson and colleagues~\citep{HassEtal08, LernEtal11, HoneEtal12, AlyEtal18}.  Subsequent work by the same group has shown that this hierarchy of drifting mental representations plays a central role in how we remember continuous experiences and segment our experiences into discrete events~\citep{BaldEtal17}. This work mirrors the discovery of \textit{time cells} in the rodent hippocampus that respond preferentially to a specific time relative to a temporal reference, whereby each cell appears to integrate incoming information at a different rate~\citep{PastEtal08, MacDEtal11} and the population activity serves to represent information at a range of timescales~\citep{MauEtal18}. Other work suggests that the lateral entorhinal cortex also plays a role in integrating information across a wide range of timescales~\citep{TsaoEtal18}.  These ideas have been formalized in a family of theoretical models over the past several decades~\citep{HowaKaha02, DianEtal07, SedeEtal08, PolyEtal09, ShanEtal09, ShanHowa10, ShanHowa12, HowaEtal14, Rang18}.  In turn, these models were inspired in part by earlier theoretical work suggesting that temporally proximal experiences become linked through their shared contextual features~\citep{Este55a,AtkiShif68}.  Collectively, these experimental and theoretical studies make two important contributions to our understanding of how we remember our past.  First, the studies suggest that our ongoing experiences are ``blurred out'' in time by neural integrators (where different brain structures or sub-structures integrate with different time constants).  Second, the studies suggest that members of this family of representations drifting at different rates become bound together such that when we retrieve memories about our past, we reactivate representations that carry information at a range of time scales.  These reactivations explain how our brains reactivate prior contexts when we remember the past.

Although reactivating context is often \textit{described} as a sort of mental time travel back to the one specific moment being remembered, there is some apparent inconsistency between this description and how this process is actually characterized mathematically or measured neurally.  In particular, both the theory and neural measures described in the preceding paragraphs characterize episodic recall (mathematically) as reflecting the reactivation of a weighted blend of thoughts from a \textit{range} of previously experienced times-- not to any one \textit{specific} time.  This is not simply a matter of imprecision in mental time travel (i.e., a reflection of the low effective resolution with which we can revisit specific moments from our past).  Rather, when we retrieve the thoughts associated with a range of times, it is conceptually more like we are simultaneously visiting \textit{many} of our prior experiences. The above human and rodent studies suggest that the brain maintains parallel representations of ongoing experience, each drifting at a different timescale.  The notion of thinking of any one particular moment from the past is incompatible with this multiple timescales view, in that our thoughts and brain activity patterns reflect information from a range of moments, even when we are not specifically engaged in remembering.  If what we consider to be a ``moment'' or ``now'' is guided by our neural representations of our ongoing experiences, then this implies we are continually experiencing \textit{many} ``nows.''  If our thoughts are continually spread across a range of times, the notion of mentally revisiting any particular moment loses its meaning.

Where the mental time travel framing breaks down most notably is when we consider scenarios in which the particular blend of timepoints reflected in one's current thoughts do not come from temporally contiguous events.  For example, understanding how a new experience fits in with our broader understanding might require integrating information gleaned from many experiences separated in time, each with their own contextual attributes.  That we can integrate information across experiences suggests that the integration of incoming information at different rates need not be a purely mechanical process, but rather may reflect a deeper functional role~\citep[also see][]{BrigEtal18}.  When we simultaneously reactivate thoughts related to separated moments from our past, this provides a means of associating and relating those temporally distinct experiences.  This is how we can integrate information across the \textit{many} discrete experiences relevant to ``now.''

@@

Event boundaries are one example of the broader temporal covariance structure that is characteristic of ongoing experience.  Whereas classic approaches to studying memory (e.g., list-learning paradigms and other trial-based experiments) encourage researchers to treat each moment, trial, or stimulus as separable from the rest of experience, context-based theories of memory (including situation models and event-based models) posit that each moment derives meaning through its deep ties to other related (but potentially temporally separated) experiences.  This rich tapestry of interacting moments and experiences forms a scaffolding for interpreting our new experiences and for retrieving information concerning our past.  According to this view, the reactivation of this rich tapestry of contextual details enfolding the remembered event (i.e., how the event relates to the rest of ones experiences across timescales) is even more central to our subjective experience than the specific details of what occurred during that event!  This begs the question: which thoughts, from which times, do we reactivate when we remember our past experiences?

\subsection*{How is our past replayed when we remember?}
Our ongoing experiences can cue us to dredge up information about our past.  This can happen explicitly (e.g., during a conversation about a specific prior event) or implicitly (e.g., when something happening now reminds you of what happened before).  Context-based theories of memory posit that the particular memories that our ongoing experiences cue will be related~\citep[contextually, semantically, functionally, etc.; e.g., ][]{PolyEtal09}.  One potential challenge to studying how we remember the past relates to detangling which links between our experiences are specifically driven by our memory systems versus which are imposed ``externally'' through the similarity structure of those experiences.  For example, consider your daily commute into work.  If you were to verbally recount a specific day's commute later, one might expect some aspects of your description to apply to other days' commutes as well (e.g., other commutes where you took a similar route or travelled via similar modes of transportation, listened to similar music along the way, travelled at a similar time, etc.).  But to what extent could one hope to separate out descriptive matches driven solely by similarities in those commuting experiences, versus true reactivations (in memory) of specific memories of those other prior commutes?

Once insight into this distinction comes from a recent study by \cite{HeusEtal18c}, which analyzed data collected by \cite{ChenEtal17} who had participants watch an episode of the BBC television show \textit{Sherlock} and then verbally recount what happened in the episode.  The authors used topic models~\citep{BleiEtal03} applied to human annotations of each camera shot in the episode to characterize the episode's content with a timeseries of high-dimensional semantic feature vectors.  These \textit{topic vectors} describe the mix of ``themes'' present in each moment of the episode.  When characterized in this way, the episode's content exhibits periods of relative stability, followed by moments of rapid change. This interplay between stability and change is visible as ``blocks'' in the temporal correlation matrix (Fig.~\ref{fig:corrmats}A); \cite{HeusEtal18c} performed detailed analyses of these putative events and how they are remembered.  Notably, the off-diagonal entries of the temporal correlation matrix (which reflect content similarity between temporally distant events in the episode) are nearly all close to 0.  This is surprising, because the episode itself comprises a murder mystery with recurring motifes that must be associated in the viewer's mind in order to make sense of the episode.  Nonetheless, when considering solely the semantic content of each moment of the episode, those long timescale associations are nearly entirely absent.  To the extent that this finding extends to real-world experiences, it comports with Heraclitus' notion that one cannot step in the same river twice~\citep{Hera}.

@@

Although the specific content of each moment of the episode is largely unique, participants' subjective experiences of the episode entail weaving together those temporally separated moments into a cohesive narrative.  This is reflected in how people verbally recount the episode later.  If we examine the temporal autocorrelation matrices of participants' recalls of the television episode (characterized using the same topic model used to capture the content of the episode; Figs.~\ref{fig:corrmats}B, C), we see a different structure than that of the episode itself.  Although individual participants' recalls exhibit periods of content stability punctuated by rapid change (block diagonal structure of the matrices in Fig.~\ref{fig:corrmats}C), the off-diagonal correlations between the thematic content of temporally distant events are substantially stronger in participants' recalls than in the original episode.  Another way of characterizing this phenomenon is to look at the autocorrelation functions (describing the correlation between different moments' topic vectors as a function of their temporal distance) for the episode and participants' recalls (Fig.~\ref{fig:reinstatement}A).  Whereas the autocorrelations in the original episode's content rapidly fall to 0, autocorrelations in the content of participants' recalls appear to asymptote around 0.2.  These analyses indicate that the way participants \textit{recount} temporally separated events imposes additional similarity on those events, beyond what was present in the original events themselves.  One can then ask: is this simply a matter of imprecision in recall, either with respect to participants' recalls themselves, or with respect to the quality with which they are characterized via semantic feature vectors?  Or might there be a functional role of that additional imposed similarity structure?

To examine which specific events are described (during recall) as more similar than what their original experiences (as characterized by their topic vectors) would suggest, it can be helpful to examine individual recalls of particular events throughout the episode.  Figure~\ref{fig:reinstatement}B displays, for a representative example recalled moment, which other moments from the episode were characterized in a similar way by the participants when they recounted the episode later (black and colored lines), and which other moments in the episode were described in a similar way in moment-by-moment (non-recall) human annotations (gray line).  The example event explored in the figure comprises a scene where Sherlock Holmes and John Watson (main characters) are bantering, and then are interrupted when Sherlock realizes that another victim has been murdered.  The language used by human annotators describing the moment-by-moment contents of the scene used a similar mix of themes only for moments in the episode that were temporally proximal to the event in question (gray line).  However, the way participants \textit{described} what happened when they recounted the episode later revealed a much different pattern.  Specifically, their language mirrored that used when they were recalling other murders, and other moments when Sherlock and John were bantering (Fig.~\ref{fig:reinstatement}C).  Further, the way they described this example event did \textit{not} mirror the way they described other semantically (but not conceptually) related events, such as events involving unrelated violence, or Sherlock and John interacting with other characters.  Taken together, this suggests that these specific events were associated in participants' memories, perhaps contributing to (or reflecting) their understanding that those events were linked in the episode's narrative.

@@

A rich recall density function analogous to the one displayed in Figure~\ref{fig:reinstatement}B could, in principle, be characteristic of memory for \textit{any} past event or experience.  For example, the peaks of the function could provide insight into which events, from which moments along one's autobiographical timeline, are functionally associated in memory.  Nonetheless, a potential challenge to testing this hypothesis in the laboratory is that in many memory studies there is no particular reason for participants to functionally associate the stimuli.  For example, there is often no ``deeper narrative'' to make sense of in list-learning studies or in most standard trial-based memory experiments.  However, random word list learning experiments might happen to contain (for a given list) semantically related words.  And indeed, a number of studies indicate that participants spontaneously form associations between semantically related words on random word lists, even when those related words are separated in time~\citep[e.g.,][]{WixtRohr94, MannKaha12, MannEtal12}.  This suggests that our memory systems pick up on statistical regularities, even in ostensibly ``random'' stimulus sequences, and that we leverage those regularities when we recall our past~\citep{PolyEtal09}.  Other work suggests that our memory systems may also leverage these regularities to parse our ongoing experiences into discrete events~\citep{SchaEtal13}.  Are these examples in the literature on random list learning studies, indicating that our memory systems leverage ``accidental'' structure in random sequences, a reflection of the same fundamental principle driving the sorts of memory reactivation function displayed in Figure~\ref{fig:reinstatement}B?

@@

\section*{The shapes of remembered experiences}
The study by \cite{HeusEtal18c} described above suggests one way of characterizing the dynamic content of complex naturalistic stimuli like television episodes.  The sequence of topic vectors over the course of the episode traces out a \textit{topic trajectory} through a high-dimensional feature space.  The sequence of topic vectors for a given participant's recalls traces out an analogous trajectory.  This characterization provides a geometric framework for comparing the \textit{shapes} that reflect how naturalistic experiences unfold over time, to the shapes of how we remember those experiences later (Fig.~\ref{fig:trajectories}).  One can then use the geometric transformations needed to map an experience's trajectory onto the trajectory of its later remembering to characterize the quality and content of memory.

@@

In addition to serving as a framework for quantifying the dynamic content of experiences, these trajectories may be visualized geometrically by projecting word embedding representations of the events onto lower dimensional spaces~\citep[e.g., ][]{HeusEtal18a}; for additional discussion and counterpoints see~\cite{JollChan18}.  \cite{HeusEtal18c} applied this approach to the same dataset reflected in Figures~\ref{fig:corrmats}, \ref{fig:reinstatement}, and~\ref{fig:trajectories}, revealing another key property of how we remember.  Although there are substantial individual differences in the specific words people use to describe a shared experience, and how much detail they get into, the overall ``shapes'' of those recalled experience trajectories are highly similar across people.  This has potentially exciting implications for how we communicate memories to other people: we need to convey the ``gist'' (shape) of our experience.  It also has implications for how we learn by analogy: perhaps events that share schemas~\citep[e.g., ][]{BaldEtal18}  unfold in similar ways (e.g., have similar shapes).  Learning which shapes go with which schemas could provide a scaffolding for rapid learning and generalization of new experiences, to the extent that those new experiences match a previously learned schema (shape).

@@

\subsection*{Extensions to spatial, semantic, and social memory}
Just as we mentally spread ourselves across a distribution of times, we can perform similar mental operations when we conceptualize and retrieve other types of information.  For example, consider how we use pattern classifiers to connect neural activity patterns to specific semantic~\citep{PolyEtal05, MitcEtal08, MannEtal12}, spatial~\citep{MillEtal13}, and social~\citep{MeyeEtal18} content being mentally manipulated.
The prevailing approach is to label specific brain activity patterns as each reflecting \textit{one particular} semantic concept or spatial location.  Variability in neural activity over repeated experiences or rememberings can result in classifiers spreading their predictions over multiple possible mental states.  Traditional approaches then ``clean up'' this distribution over possible mental states by taking its expected or maximum value.  However, following the ``quantum wave function'' logic described in this manuscript, I suggest that the specific shape of these distributions may not only be informative, but they are \textit{more accurate reflections of what our neural activity patterns actually reflect}.  What are the implications of this view?

\begin{itemize}
\item event boundaries and goal boundaries-- environment barriers, non-sensical concepts \item not every nearby stimulus is equally likely or equally reachable.  Stimuli live on a manifold; could use the shape to understand the geometry of thought, nature of problem space, improve decoding, etc.
  \item links to GANs (pix2pix, etc.)?
\end{itemize}

\section*{Concluding remarks}
When we revisit an experience from our past, we bring back thoughts related to a range of other semantically and functionally related experiences.  This can serve an important functional role: it allows us to integrate information across temporally separated experiences.  Although this view is captured mathematically by context-based theories of episodic memory, it is \textit{not} specifically captured by the conceptual notion of mental time travel that is often used to describe episodic recall.  The alternative conceptualization presented here, that when we remember our past we distribute our thoughts over those associated with a weighted blend of our prior experiences, also has links with non-temporal or episodic domains such as semantic, spatial, and social representations.  Characterizing how our thoughts reflect \textit{distributions} over representational spaces is central to understanding how we think about ourselves, our experiences, other people, and the world around us.
